Quarto Template for PNAS Submissions

### 2023-12-02 manuscript/revision2/revision2.tex
Add revision Quarto document in PNAS format

---
title: Quarto Template for PNAS Submissions
format:
  pnas-pdf:
    keep-tex: true  
author:
  - name: Author One
    affiliations:
      - name: Harvard University
        id: a
        department: Department of Government
        city: Cambridge
        state: MA
        postal-code: 02138
      - name: Harvard University
        id: c
        department: Department of Statistics
        city: Cambridge
        state: MA
        postal-code: 02138
    attributes:
      equal-contributor: true
  - name: Author Two
    affiliations:
      - name: Yale University
        id: b
        department: Department of Political Science
        city: New Haven
        state: CT
        postal-code: 06511
    email: alicem@uchicago.edu
    attributes:
      equal-contributor: true
      corresponding: true
  - name: Author Three
    affiliations:
      - ref: a
abstract: |
  The timing of vegetative growth and flowering in plants ("phenological timings") multiple depend on environmental cues; thus, phenological traits can have important genotype-by-environment (GxE) interactions. Here, we map genotype-by-weather effects on phenological timings in multiple populations of two highly divergent switchgrass (Panicum virgatum) populations spanning the central United States. We distinguish GxWeather patterns covarying with interpretable weather-based cues and agnostic, empirically-inferred GxWeather patterns. Most GxWeather effects belong to the latter category and do not covary with weather-based cues. However, >26% of Gulf population effects on flowering time covary with photoperiod cues, while >__% of Midwest upland population SNPs affecting flowering had effects that covaried with __ cues. An independent pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs for flowering at the same common gardens, all with significant mash associations and ten with enrichment of highly significant mash associations. We demonstrate that we can identify QTL with GxE and assign them to specific weather-based cues and/or data-driven patterns. Breeding for particular alleles at these loci could change flowering responsiveness to photoperiod cues in switchgrass. More broadly, this approach could be used to identify genetic marker-environment interactions in any species with related populations phenotyped in multiple environments.
runningauthor: "One, Two, and Three"
significance: |
  The timing of plant seasonal development (phenology) has major impacts on fitness because of the steep price of plant-environment mismatches. Here we map the genetic basis of two phenological events, the start of above-ground growth and flowering, in two genetically and phenologically distinct populations of switchgrass. We do this at eight field locations spanning the latitudinal range of both switchgrass populations. Our approach allows us to identify regions of the genome that covary with specific weather-related environmental features at every location. For flowering, these features differed by population: the Midwest population had genetic variation that primarily covaried with cumulative growing degree days, a temperature-related measure, while the Gulf population had genetic variation that primarily covaried with photoperiod, a day-length-related measure.
author_contributions:  "T.E.J. designed research. D.B.L. contributed plant material and resources. J.B., D.B.L., and T.E.J. designed and executed field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W., and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted statistical and computational analyses. The manuscript was written by A.H.M., L.Z., and T.E.J. with contributions from all authors."
conflict_of_interest: "Please declare any competing interests here."
equal-contributor: "A.O.(Author One) contributed equally to this work with A.T. (Author Two) (remove if not applicable)."
keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative growing degree days, genetic variation]
acknowledgements: | 
  We thank the Brackenridge Field laboratory, the Ladybird Johnson Wildflower Center, and the Juenger laboratory for support with plant care and propagation. This material is based upon work supported in part by the Great Lakes Bioenergy Research Center, U.S. Department of Energy, Office of Science, Office of Biological and Environmental Research under Award Numbers DE-SC0018409 and DE-FC02-07ER64494, the US Department of Energy Awards DESC0014156 to T.E.J., DE-SC0017883 to D.B.L, National Science Foundation PGRP Awards IOS0922457 and IOS1444533 to T.E.J, and the Long-term Ecological Research Program (DEB 1832042) at the Kellogg Biological Station.
bibliography: bibliography.bib  
fontsize: "9pt"
classoption:
 - twocolumn
 - twoside
---

\firstpage[20]{2}

The timing of plant vegetative and reproductive development are major components of plant fitness affected by multiple external environmental cues (e.g. degree of winter chilling, day length, temperature, and water availability) that signal existing or upcoming growing conditions (1--3). Genetic responses to environmental cues determine the speed, timing, and energy apportioned to vegetative and reproductive growth and shape both the individual's lifespan and its lifetime production of viable seed. Day length (or photoperiod) is one of the most predictable environmental cues, and genetic sensitivity to photoperiod protects plants from potentially fatal consequences of phenological responses to temperature cues at the "wrong" time of year. However, the usefulness of specific environmental cues depends on both features of the environment and the species' adaptive strategies (4). Species with wide natural distributions can have multiple distinct environmentally-cued phenological responses: for example, populations of sunflower (Helianthus annuus) exhibit day-neutral, facultative short day, and facultative long-day flowering responses, which vary with their environments (5, 6). Distinct genetic responses in different environments are known as genotype by environment interactions, or GxE.

@@

Here, we test how the Midwest and Gulf subpopulations differ in their phenological adaptations across environments, and hence their phenological GxE. We phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative development ("green-up") and reproductive development (flowering) at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. After single-site estimation of SNP effects for these phenotypes, we use multivariate adaptive shrinkage (mash) to jointly estimate SNP effects in each subpopulation and phenotype at all eight sites (35). Mash allows us to identify and specify multiple ways genetic effects linked to SNPs may covary with the environment, and does not have a statistical bias in detecting SNP effects with the same or opposite signs (39). To confirm our genetic mapping of GxE, we compare to mapping results from an outbred pseudo-F2 cross grown at the same sites. Taken together, our results allow us to describe the environmental cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.

# Results

In our diversity panel of tetraploid switchgrass (32), genotypes from the Gulf and Midwest genetic subpopulations had distinct phenological timings and distinct patterns of phenological correlations across our eight common garden sites (Fig. 1A,B). At the three Texas common gardens (hereafter 'Texas' gardens), located within the natural range of the Gulf subpopulation, Gulf green-up occurred before Midwestern green-up, and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the four northernmost common gardens (hereafter 'North' gardens), located within the natural range of the Midwest subpopulation, both Gulf green-up and flowering occurred after Midwest green-up and flowering. At the Oklahoma common garden, located near the natural range limits of both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up occurred over the same time period. These patterns led to strong negative phenotypic correlations for green-up between the North and Texas gardens, particularly in the Gulf and Both subpopulations, and contributed to positive phenotypic correlations for flowering time of larger magnitude at more northern gardens (Fig. 1B).

```         
Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C). h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date. However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 < 0.2) or negatively correlated between pairs of gardens (Fig. 1B). These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
```

### Mapping major patterns of genotype-by-environment effects on green-up and flowering

```         
We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype. We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices. 
```

Three types of covariance matrices were tested for inclusion: "canonical" covariance matrices, with simple patterns of effects; "data-driven" matrices derived from common patterns of SNP effects in the data, and "hypothesis-based" covariance matrices that captured the covariance of weather-based phenological cues for cloned genotypes grown at multiple common gardens (Table 1, SI Appendix, Section S2). Mash first assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood. Then, in a second step, each SNP's effect estimate undergoes shrinkage such that it is more aligned with the covariance structure inferred in the first step. The hypothesis-based covariance matrices are an important advantage mash offers for studying patterns of GxE: they allow hypothesis testing of specific environmental drivers for each SNP effect, and loadings on these matrices provide information on genome-wide patterns of SNP-environment interaction. Say that a SNP in FLC controls flowering in a temperature-dependent manner. In that case, that SNP effect could have a high mixture proportion, or mass, on a covariance matrix created using a temperature-based environmental cue, such as average temperature on the day prior to flowering. In our data, we would infer that the effect of that SNP on flowering was caused by a response to that or a correlated environmental cue.

The phenotypic correlations for green up date had moderate negative correlations between the Texas and North gardens, particularly in the Gulf and in Both subpopulations (Fig 1B). If these phenotypic correlations have a genetic component, then they might be controlled by SNP effects that differed in sign across these regions, SNP effects that are large in one garden or region and non-significant in others, or a combination of these patterns. Five hypothesis-based covariance matrices were selected by the greedy mash algorithm, one to two per subpopulation. Of these five, three had mass on them in mash models of the strong effects (Fig 2A, 2B). Two of these three matrices had negative covariances between sites between Texas and North gardens (Fig 2A), while one had all positive or near-zero covariances. SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (Fig 2B). In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix of daylength 14 days prior to the date of green-up. The covariance matrix for this weather cue was very similar to the pattern of phenotypic correlation for green-up in the Gulf subpopulation (Fig 1B; Fig 2A). Mash models of Midwest and Both subpopulation green-up did not include this covariance matrix; the Midwest had no weight on any hypothesis-based matrix, while Both subpopulations had non-zero weights on two additional hypothesis-based matrices, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (Fig 2B). The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (Fig 1B; Fig 2A). Only the Gulf subpopulation and Both subpopulations had mass on any hypothesis-based covariance matrices (Fig 2C).

For flowering date, distinct weather-based cues captured SNP-associated effect patterns in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (Fig 2D). 22.6% of SNP effects on flowering in the Midwest subpopulation contrived with day length change in the two days prior to flowering (Fig 2D), which had negative covariances between Texas and North gardens. Neither covariance matrix was selected for in Both subpopulations, and no hypothesis-based covariance matrices had mass in Both subpopulations (Fig 2E). In five of the six mash models of strong effects, the hypothesis-based covariance matrices captured a minority of the posterior weights of the strong effects (Figure 2C,E); the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens. We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance. We required significance (p \< 0.05) in both gardens to include effects. This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign. We also summarized the pairwise patterns of effect sign and magnitude within and between the Texas and North regions (Table 2).

For greenup date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (Table 2). The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects frequently differed in magnitude between the MO and OK garden and other gardens (Table 2; Fig 3A).

For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (Table 2). More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; Figure 3B). The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%). Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (Table 2). Most differences in sign were between TX1, the southernmost garden, and all other gardens (Fig 3B). Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).

### Confirmation of genotype-by-environment effects using an independent mapping population

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  9pt,
  twocolumn,
  twoside]{pnas-new}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% TODO: Add custom LaTeX header directives here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Quarto Template for PNAS Submissions},
  pdfauthor={Author One; Author Two; Author Three},
  pdfkeywords={allele-by-environment effect variation, antagonistic
pleiotropy, photoperiod, cumulative growing degree days, genetic
variation},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\templatetype{pnasresearcharticle}

\title{Quarto Template for PNAS Submissions}

\newcommand{\equalcont}{1}

\newcommand{\correspond}{2}

\author[a,c%
,\equalcont%
%
]{Author One}
\author[b%
,\equalcont%
,\correspond%
]{Author Two}
\author[a%
%
%
]{Author Three}

\affil[a]{Harvard University, Department of Government, Cambridge, 2138}
\affil[c]{Harvard University, Department of Statistics, Cambridge, 2138}
\affil[b]{Yale University, Department of Political Science, New
Haven, 3401}

\leadauthor{One, Two, and Three}

\significancestatement{The timing of plant seasonal development
(phenology) has major impacts on fitness because of the steep price of
plant-environment mismatches. Here we map the genetic basis of two
phenological events, the start of above-ground growth and flowering, in
two genetically and phenologically distinct populations of switchgrass.
We do this at eight field locations spanning the latitudinal range of
both switchgrass populations. Our approach allows us to identify regions
of the genome that covary with specific weather-related environmental
features at every location. For flowering, these features differed by
population: the Midwest population had genetic variation that primarily
covaried with cumulative growing degree days, a temperature-related
measure, while the Gulf population had genetic variation that primarily
covaried with photoperiod, a day-length-related measure.}

\authorcontributions{T.E.J. designed research. D.B.L. contributed plant
material and resources. J.B., D.B.L., and T.E.J. designed and executed
field experiments. A.R.B., P.A.F., F.B.F., D.B.L., R.B.M., F.M.R., Y.W.,
and T.E.J. hosted field experiments. A.H.M. and L.Z. conducted
statistical and computational analyses. The manuscript was written by
A.H.M., L.Z., and T.E.J. with contributions from all authors.}

\authordeclaration{Please declare any competing interests here.}

\equalauthors{\textsuperscript{\equalcont} A.O.(Author One) contributed
equally to this work with A.T. (Author Two) (remove if not applicable).}

\correspondingauthor{\textsuperscript{\correspond}To whom correspondence should be addressed. E-mail: alicem@uchicago.edu}

\keywords{allele-by-environment effect variation | antagonistic
pleiotropy | photoperiod | cumulative growing degree days | genetic
variation}

\begin{document}
\maketitle

\begin{abstract}
The timing of vegetative growth and flowering in plants (``phenological
timings'') multiple depend on environmental cues; thus, phenological
traits can have important genotype-by-environment (GxE) interactions.
Here, we map genotype-by-weather effects on phenological timings in
multiple populations of two highly divergent switchgrass (Panicum
virgatum) populations spanning the central United States. We distinguish
GxWeather patterns covarying with interpretable weather-based cues and
agnostic, empirically-inferred GxWeather patterns. Most GxWeather
effects belong to the latter category and do not covary with
weather-based cues. However, \textgreater26\% of Gulf population effects
on flowering time covary with photoperiod cues, while
\textgreater{}\textbf{\% of Midwest upland population SNPs affecting
flowering had effects that covaried with } cues. An independent
pseudo-F2 cross of Gulf and Midwest individuals mapped 23 additive QTLs
for flowering at the same common gardens, all with significant mash
associations and ten with enrichment of highly significant mash
associations. We demonstrate that we can identify QTL with GxE and
assign them to specific weather-based cues and/or data-driven patterns.
Breeding for particular alleles at these loci could change flowering
responsiveness to photoperiod cues in switchgrass. More broadly, this
approach could be used to identify genetic marker-environment
interactions in any species with related populations phenotyped in
multiple environments.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}\firstpage[20]{2}

The timing of plant vegetative and reproductive development are major
components of plant fitness affected by multiple external environmental
cues (e.g.~degree of winter chilling, day length, temperature, and water
availability) that signal existing or upcoming growing conditions
(1--3). Genetic responses to environmental cues determine the speed,
timing, and energy apportioned to vegetative and reproductive growth and
shape both the individual's lifespan and its lifetime production of
viable seed. Day length (or photoperiod) is one of the most predictable
environmental cues, and genetic sensitivity to photoperiod protects
plants from potentially fatal consequences of phenological responses to
temperature cues at the ``wrong'' time of year. However, the usefulness
of specific environmental cues depends on both features of the
environment and the species' adaptive strategies (4). Species with wide
natural distributions can have multiple distinct environmentally-cued
phenological responses: for example, populations of sunflower
(Helianthus annuus) exhibit day-neutral, facultative short day, and
facultative long-day flowering responses, which vary with their
environments (5, 6). Distinct genetic responses in different
environments are known as genotype by environment interactions, or GxE.

@@

Here, we test how the Midwest and Gulf subpopulations differ in their
phenological adaptations across environments, and hence their
phenological GxE. We phenotype a diversity panel of hundreds of
switchgrass genotypes from the Midwest and Gulf subpopulations for the
timing of vegetative development (``green-up'') and reproductive
development (flowering) at eight common garden locations spanning 17
degrees of latitude. These gardens cover the majority of the latitudinal
and climatic range of switchgrass and capture the most comprehensive
picture to date of the environmental variation this species encounters.
After single-site estimation of SNP effects for these phenotypes, we use
multivariate adaptive shrinkage (mash) to jointly estimate SNP effects
in each subpopulation and phenotype at all eight sites (35). Mash allows
us to identify and specify multiple ways genetic effects linked to SNPs
may covary with the environment, and does not have a statistical bias in
detecting SNP effects with the same or opposite signs (39). To confirm
our genetic mapping of GxE, we compare to mapping results from an
outbred pseudo-F2 cross grown at the same sites. Taken together, our
results allow us to describe the environmental cues and genetic
variation affecting phenology in two divergent natural populations of
switchgrass.

\section{Results}\label{results}

In our diversity panel of tetraploid switchgrass (32), genotypes from
the Gulf and Midwest genetic subpopulations had distinct phenological
timings and distinct patterns of phenological correlations across our
eight common garden sites (Fig. 1A,B). At the three Texas common gardens
(hereafter `Texas' gardens), located within the natural range of the
Gulf subpopulation, Gulf green-up occurred before Midwestern green-up,
and Gulf flowering occurred after Midwestern flowering (Fig. 1A). At the
four northernmost common gardens (hereafter `North' gardens), located
within the natural range of the Midwest subpopulation, both Gulf
green-up and flowering occurred after Midwest green-up and flowering. At
the Oklahoma common garden, located near the natural range limits of
both the Gulf and the Midwest subpopulations, Gulf and Midwest green-up
occurred over the same time period. These patterns led to strong
negative phenotypic correlations for green-up between the North and
Texas gardens, particularly in the Gulf and Both subpopulations, and
contributed to positive phenotypic correlations for flowering time of
larger magnitude at more northern gardens (Fig. 1B).

\begin{verbatim}
Narrow-sense heritabilities (h2) indicated that rank-changing GxE for these phenotypes was present across the common gardens (Fig 1C). h2 were typically high at individual gardens: 59% on average for green-up date, and 87% for flowering date. However, h2 were variable across gardens, and green-up dates were uncorrelated (r2 < 0.2) or negatively correlated between pairs of gardens (Fig. 1B). These negative and small correlations undoubtedly contributed to the low h2 values for green-up and flowering date when estimated jointly at all eight gardens: h2 was 0.8% for green-up and 23.2% for flowering date.
\end{verbatim}

\subsubsection{Mapping major patterns of genotype-by-environment effects
on green-up and
flowering}\label{mapping-major-patterns-of-genotype-by-environment-effects-on-green-up-and-flowering}

\begin{verbatim}
We next explored how genetic variation in phenology covaried with environmental cues by using mash to jointly discover trait-associated SNPs across all common gardens for the Midwest, Gulf, and ‘Both’ subpopulations (Fig 1D, SI Appendix, Datasets 1-6). We first estimated SNP effects using GWAS on site-specific BLUPs for each phenotype. We then used mash on two subsets of these SNP effects, first to determine the covariance structures that significantly improved the mash model log likelihood when included, then to re-estimate the set of SNP effects with the largest effect estimates in univariate GWAS (hereafter “strong” effects) using the set of significant covariance matrices. 
\end{verbatim}

Three types of covariance matrices were tested for inclusion:
``canonical'' covariance matrices, with simple patterns of effects;
``data-driven'' matrices derived from common patterns of SNP effects in
the data, and ``hypothesis-based'' covariance matrices that captured the
covariance of weather-based phenological cues for cloned genotypes grown
at multiple common gardens (Table 1, SI Appendix, Section S2). Mash
first assigns mixture proportions for each SNP onto each provided
covariance matrix using maximum likelihood. Then, in a second step, each
SNP's effect estimate undergoes shrinkage such that it is more aligned
with the covariance structure inferred in the first step. The
hypothesis-based covariance matrices are an important advantage mash
offers for studying patterns of GxE: they allow hypothesis testing of
specific environmental drivers for each SNP effect, and loadings on
these matrices provide information on genome-wide patterns of
SNP-environment interaction. Say that a SNP in FLC controls flowering in
a temperature-dependent manner. In that case, that SNP effect could have
a high mixture proportion, or mass, on a covariance matrix created using
a temperature-based environmental cue, such as average temperature on
the day prior to flowering. In our data, we would infer that the effect
of that SNP on flowering was caused by a response to that or a
correlated environmental cue.

The phenotypic correlations for green up date had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf and in Both subpopulations (Fig 1B). If these phenotypic
correlations have a genetic component, then they might be controlled by
SNP effects that differed in sign across these regions, SNP effects that
are large in one garden or region and non-significant in others, or a
combination of these patterns. Five hypothesis-based covariance matrices
were selected by the greedy mash algorithm, one to two per
subpopulation. Of these five, three had mass on them in mash models of
the strong effects (Fig 2A, 2B). Two of these three matrices had
negative covariances between sites between Texas and North gardens (Fig
2A), while one had all positive or near-zero covariances. SNP-associated
phenotypic effects covaried with different weather-based cues in the
Gulf \& in Both subpopulations (Fig 2B). In total, 65\% of the posterior
weight of strong SNP effects in the mash model of Gulf green-up fell on
a covariance matrix of daylength 14 days prior to the date of green-up.
The covariance matrix for this weather cue was very similar to the
pattern of phenotypic correlation for green-up in the Gulf subpopulation
(Fig 1B; Fig 2A). Mash models of Midwest and Both subpopulation green-up
did not include this covariance matrix; the Midwest had no weight on any
hypothesis-based matrix, while Both subpopulations had non-zero weights
on two additional hypothesis-based matrices, average temperature one day
prior to green-up, and the day length change in seconds in the day prior
to green-up (Fig 2B). The average temperature covariance matrix had
negative covariances between Texas and North gardens, though not as
strong as the negative phenotypic correlations seen in Both
subpopulations (Fig 1B; Fig 2A). Only the Gulf subpopulation and Both
subpopulations had mass on any hypothesis-based covariance matrices (Fig
2C).

For flowering date, distinct weather-based cues captured SNP-associated
effect patterns in the Gulf and Midwest subpopulations. 33\% of SNP
effects on flowering in the Gulf subpopulation covaried with cumulative
rainfall in the seven days prior to flowering (Fig 2D). 22.6\% of SNP
effects on flowering in the Midwest subpopulation contrived with day
length change in the two days prior to flowering (Fig 2D), which had
negative covariances between Texas and North gardens. Neither covariance
matrix was selected for in Both subpopulations, and no hypothesis-based
covariance matrices had mass in Both subpopulations (Fig 2E). In five of
the six mash models of strong effects, the hypothesis-based covariance
matrices captured a minority of the posterior weights of the strong
effects (Figure 2C,E); the majority of this mass was on various
canonical covariance matrices. These matrices included simple
heterozygosity, with intermediate, positive covariances between all
gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were
confident in the sign of the effect at both gardens. We used the local
false sign rate (lfsr), an analogue of the lfdr that establishes
confidence in the effect sign, not the effect's difference from zero, to
determine significance. We required significance (p \textless{} 0.05) in
both gardens to include effects. This means that our tests for
antagonistic pleiotropy, or a sign change between conditions, carry an
equal statistical burden to those for effects with the same sign. We
also summarized the pairwise patterns of effect sign and magnitude
within and between the Texas and North regions (Table 2).

For greenup date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited antagonistic pleiotropy, or a difference in
effect sign, between pairs of Texas and North gardens (Fig. 3A). 78.7\%
of pairwise comparisons between North and Texas gardens had a difference
in sign, while only 28.6\% and 0.2\% of North-North or Texas-Texas
comparisons had a difference in sign, respectively (Table 2). The
majority of pairwise effects for greenup for the Midwest
(\textgreater55\%) and Both (\textgreater85\%) subpopulations were the
same sign, and effects frequently differed in magnitude between the MO
and OK garden and other gardens (Table 2; Fig 3A).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited antagonistic pleiotropy, or a difference in effect
sign, within or between regions (Table 2). More effects differed in
magnitude between the Texas and North regions than within these regions
(42.7\% vs \textless20\%; Figure 3B). The Midwest population had
relatively few significant effects for flowering, but a large proportion
of these differed in sign between Texas and North regions (42.7\%) or
within the North region (65.4\%). Finally, in Both subpopulations, less
than 20\% of pairwise effects differed in sign (Table 2). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (Fig 3B). Similarly, more effects that differed in
magnitude included gardens in the Texas region (52.3-55.9\%), and most
effect pairs in the North region were not distinguishable (91.5\%).

\subsubsection{Confirmation of genotype-by-environment effects using an
independent mapping
population}\label{confirmation-of-genotype-by-environment-effects-using-an-independent-mapping-population}
